Analyzing Fairness: Race, Recidivism, and the COMPAS Algorithm

### 2022-04-28 Markdown Files Sprints 3 and 4/final_metrics_analysis.Rmd

# for Dianne
# use_condaenv("base")

@@

Pre-processing refers to mitigating bias within the training data. For our pre-processing approach we are using AIF360's Reweighing algorithm. The reweighing approach "assigns different weights to the examples based upon their categories of protected attribute and outcome such that bias is removed from the training dataset" ( [Mahmoudian](https://towardsdatascience.com/reweighing-the-adult-dataset-to-make-it-discrimination-free-44668c9379e8)).

## Reweighing

@@

In-processing mitigates bias in classifiers while building a model. The in-processing technique we use is the prejudice remover algorithm from AIF360. The prejudice remover is a method for reducing indirect prejudice (i.e. how the COMPAS algorithm is racially biased because it uses proxy variables for race). This approach implements two different regularizers, one to avoid overfitting and one to enforce fair classification ( [Kamishima et. al.](https://link.springer.com/content/pdf/10.1007/978-3-642-33486-3_3.pdf)). The prejudice remover regularizer works by minimizing the prejudice index, a mathematical equation for quantifying fairness defined by [Kamishima et. al](https://link.springer.com/content/pdf/10.1007/978-3-642-33486-3_3.pdf). This in turn enforces a classifier’s independence from sensitive information (like race).

@@

Our last approach, post-processing bias mitigation, is used after training a model. Post-processing algorithms equalize the outcomes (i.e., predicted classification values) to mitigate bias instead of adjusting the classifier or the training data( [Baxter](https://www.infoworld.com/article/3630450/what-is-ai-bias-mitigation-and-how-can-it-improve-ai-fairness.html)). We use calibrated equalized odds, which "optimizes over calibrated classifier score outputs to find probabilities with which to change output labels with an equalized odds objective" ( [AIF360](https://aif360.readthedocs.io/en/stable/modules/generated/aif360.algorithms.postprocessing.CalibratedEqOddsPostprocessing.html#aif360.algorithms.postprocessing.CalibratedEqOddsPostprocessing)). An equalized odds objective constrains classification algorithms such that no error type (false-positive or false-negative) disproportionately affects any population subgroup; both groups, in our case both white and Black defendants, should have the same false-positive and false-negative rates. Through the calibrated equalized odds method, we want to decrease bias while also maintaining calibration ( [Pleiss et. al.](https://arxiv.org/pdf/1709.02012.pdf)). [Calibration](https://medium.com/analytics-vidhya/calibration-in-machine-learning-e7972ac93555) refers to improving a model so that the distribution of predicted outcomes is similar to the distribution of observed probability in the training data.

## Calibrated Equalized Odds

>>>>

# for Dianne
use_condaenv("base")

@@

## Reweighting

@@

@@
